---
title: "Case Study 1"
author: "Chris Irwin"
date: "11/1/2016"
output: html_document
---
#Analysis Instructions:
In order to start the analysis please start by opening r and set the working directory of your choice, using the setwd("Your Location goes here"). It is very important that after setting your working directory you will need to create a folder called Analysis. Once the folder is created then download all the files located in the Analysis folder inside of GITHub. After starting the execution of the r script it is important to know that the script will create one new folder called Data. Inside of this folder will save the files after being downloaded from the URLs above. These files are created in case you wanted to take the analysis offline and not have to rely on re-downloading the file each time. 


#Purpose:
This case study consists of 2 files, which are located on the internet and need to combined in order to answer the following questions.

1. Merge the data based on the country shortcode. How many of the IDs match?

2. Sort the data frame in ascending order by GDP (so United States is last). What is the 13th
  country in the resulting data frame?

3. What are the average GDP rankings for the "High income: OECD" and "High income:
  nonOECD" groups?

4. Plot the GDP for all of the countries. Use ggplot2 to color your plot by Income Group.

5. Cut the GDP ranking into 5 separate quantile groups. Make a table versus Income.Group.
How many countries are Lower middle income but among the 38 nations with highest GDP?

In order to answer these questions we need to download the files from the internet using the following locations. 

  1. https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FGDP.csv 
      a. This CSV file contains the GDP Data for 190 Countries 
  2. https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FEDSTATS_Country.csv 	
	    b. This csv file contains additional information for 234 countries

In order to complete the requested analysis we will need to download the listed files above using r and then merge the data sets using the country code, which is provided in both files. 


###Set Working Directory 
```{r}
setwd("/Users/cmirwin/Desktop/SMU Data Sets/Doing Data Science/Case Study 1")
```

###Create Folder Structures
```{r results = 'hide'}
source("./Analysis/CreateFolderStructure.txt")
```

###Download GDP File from https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FGDP.csv
```{r}
source("./Analysis/DownloadGDPFile.txt")
```

###Download GDP File from https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FEDSTATS_Country.csv
```{r}
source("./Analysis/DownloadFedStatsFile.txt")
```

#Results
###Merging two data files into single data frame
```{r}
source("./Analysis/MergeData.txt")
```

##Question 1
####Get the length of the matched datasets
```{r}
length(df.MergedData$CountryCode)
```

##Question 2
####Country with the 13th lowest GDP
```{r}
source("./Analysis/Question2.txt")

#Select the Short Name from the 13th GDP 
GDP.Info[c("Short.Name")]
```
##Question 3
####average GDP rankings for the "High income: OECD" and "High income:nonOECD" groups
```{r}
source("./Analysis/Question3.txt")

#Average Rank per Income Group
average.RankIncomeGroup
```


##Question 4
####Execute and Run the Entire Graph
```{r}
#Execute Script to build Graph
source("./Analysis/Question4.txt")

#Create the Graph
plot + point + title + labels + legendInfo
```

##Question 5
####Create table of Income Group Vs Quantiles
```{r}
source("./Analysis/Question5.txt")

Question5Table[,Select = (">80%")]
```

#Conclusion:
After cleaning the GDP file we find that there are 189 matching countries that exist in both the GDP and Fed Stats file. While looking at the merged data set I sorted the data by GDP and found that the 13th country in the data set is St. Kitts and Nevis. Additionally after looking into the data I found the average GDP rankings between the two High income groups, OECD and Non-OECD had a difference of 60 places between the groups. After digging in a little further I found that the OECD group average was 32.96, while the Non-OECD's average is 91.91.  Once I saw the large difference in the rankings I wanted to see how the groups faired against each other in a single graph based on overall GDP. After plotting the data I found a very interesting outlier which was India that had a very large GDP and being considered Lower Middle Income nation. Finally I broke the data into 5 groups (<20%,21%-40%,41%-60%,61%-80%,>80%), which was based on GDP of the country. Once the groups were applied I saw there there is quite an interesting fact, which is that there are 5 countries that are in the top 80% of total GDP and considered Lower Middle Income. 